\Large Clustering Vehicle Trips Identified from Automatic \\ Number Plate Recognition Camera Scans

### 2018-01-31 classification.tex
Added conclusion and future work

Tables~\ref{t:kmeans_centers_450} and~\ref{t:kmeans_centers_1200} depict the cluster centers for $\Tau$ equal to 7.5 minutes and 20 minutes respectively. The values for the cluster centers partially meet our expectations. We were expecting to find a relatively small cluster representing taxis with a high average number of trips per day, occurring over a variety of origins and destinations and over a large time frame. Clusters 2 and 7 for $\Tau = 7.5$ and cluster 4 for $\Tau = 20$ do indeed fit this profile. What differentiates cluster 2 from 7 in the first case is essentially the mean number of trips per day and the average time at which the first and last trips of the day occur. We could expect that buses could be separated from taxis by showing less diversity in the number of origins and destinations as these essentially do multiple runs of the same trip throughout the day. Still, this can be explained by the fact that buses take routes through main and secundary roads. As most cameras are placed in main roads, the one long bus trip can be perceived as multiple small trips as the bus alternates between arterial and main roads.

On the other hand, we could expect that commuters display average and relatively few trips per day.

\begin{table*}[t]
\centering
\small
\begin{tabular}{c c c c c c c c c c}
  \hline
 Cluster &  Size & \thead{Total\\Trips} & \thead{Average\\Trips} & \thead{Average\\Length} & \thead{Average\\Distinct\\Origins} & \thead{Average\\Distinct\\Destinations} & \thead{Average\\First\\Hour} & \thead{Average\\Last\\Hour} & \thead{Average\\Rest\\Time} \\
  \hline
  1 & 108868 & 30.67 & 2.91 & 1.56 & 2.65 & 1.09 & 9.67 & 16.11 & 6.35 \\
  2 & 2948 & 170.08 & 13.69 & 1.47 & 9.20 & 4.67 & 6.89 & 18.69 & 11.65 \\
  3 & 309748 & 5.95 & 2.01 & 1.44 & 1.92 & 0.68 & 12.52 & 14.50 & 1.93 \\
  4 & 163982 & 16.99 & 2.30 & 1.46 & 2.13 & 0.77 & 11.12 & 14.99 & 3.81 \\
  5 & 45414 & 50.30 & 4.26 & 1.52 & 3.67 & 1.57 & 9.04 & 16.86 & 7.69 \\
  6 & 10380 & 86.99 & 7.16 & 1.43 & 5.55 & 2.49 & 8.41 & 17.60 & 9.03 \\
  7 & 666 & 331.50 & 26.43 & 1.57 & 9.31 & 5.17 & 5.73 & 19.59 & 13.14 \\
   \hline
\end{tabular}
\caption{Clusters sizes and mean centers for $\Tau = 7.5$ minutes.}
\label{t:kmeans_centers_450}
\end{table*}

\begin{table*}[t]
\centering
\small
\begin{tabular}{c c c c c c c c c c}
  \hline
 Cluster &  Size & \thead{Total\\Trips} & \thead{Average\\Trips} & \thead{Average\\Length} & \thead{Average\\Distinct\\Origins} & \thead{Average\\Distinct\\Destinations} & \thead{Average\\First\\Hour} & \thead{Average\\Last\\Hour} & \thead{Average\\Rest\\Time} \\
  \hline
  1 & 371318 & 7.19 & 1.76 & 1.68 & 1.68 & 0.71 & 12.30 & 14.55 & 2.14 \\
    2 & 49852 & 44.85 & 3.73 & 1.90 & 3.16 & 1.73 & 8.98 & 17.02 & 7.76 \\
    3 & 189224 & 22.82 & 2.29 & 1.84 & 2.11 & 1.06 & 10.00 & 15.76 & 5.60 \\
    4 & 4593 & 114.34 & 9.12 & 2.30 & 5.92 & 3.98 & 6.93 & 18.21 & 10.43 \\
   \hline
\end{tabular}
\caption{Clusters sizes and mean centers for $\Tau = 20$ minutes.}
\label{t:kmeans_centers_1200}
\end{table*}

>>>>

Tables~\ref{t:kmeans_centers_450} and~\ref{t:kmeans_centers_1200} depict the cluster centers for $\Tau$ equal to 7.5 minutes and 20 minutes respectively. The values for the cluster centers partially meet our expectations. We were expecting to find a relatively small cluster representing taxis with a high average number of trips per day, occurring over a variety of origins and destinations and over a large time frame. Clusters 2 and 7 for $\Tau = 7.5$ and cluster 4 for $\Tau = 20$ do indeed fit this profile. What differentiates cluster 2 from 7 in the first case is essentially the mean number of trips per day and the average time at which the first and last trips of the day occur. We expected that buses could be separated from taxis by showing less diversity in the number of origins and destinations as these essentially do multiple runs of the same trip throughout the day. Still, this can be explained by the fact that buses take routes through main and secundary roads. As most cameras are placed in main roads, the one long bus trip can be perceived as multiple small trips as the bus alternates between arterial and main roads.

On the other hand, we could expect that commuters display trips starting at around 8h in the morning and terminating at 17h-18h. Additionally, we would expect to see few trips per day as most of the day is spent in work, ocasionally .

\begin{table*}[t]
\centering
\small
\begin{tabular}{c c c c c c c c c c}
  \hline
 Cluster &  Size & \thead{Total\\Trips} & \thead{Average\\Trips} & \thead{Average\\Length} & \thead{Average\\Distinct\\Origins} & \thead{Average\\Distinct\\Destinations} & \thead{Average\\First\\Hour} & \thead{Average\\Last\\Hour} & \thead{Average\\Rest\\Time} \\
  \hline
  1 & 108868 & 30.67 & 2.91 & 1.56 & 2.65 & 1.09 & 9.67 & 16.11 & 6.35 \\
  2 & 2948 & 170.08 & 13.69 & 1.47 & 9.20 & 4.67 & 6.89 & 18.69 & 11.65 \\
  3 & 309748 & 5.95 & 2.01 & 1.44 & 1.92 & 0.68 & 12.52 & 14.50 & 1.93 \\
  4 & 163982 & 16.99 & 2.30 & 1.46 & 2.13 & 0.77 & 11.12 & 14.99 & 3.81 \\
  5 & 45414 & 50.30 & 4.26 & 1.52 & 3.67 & 1.57 & 9.04 & 16.86 & 7.69 \\
  6 & 10380 & 86.99 & 7.16 & 1.43 & 5.55 & 2.49 & 8.41 & 17.60 & 9.03 \\
  7 & 666 & 331.50 & 26.43 & 1.57 & 9.31 & 5.17 & 5.73 & 19.59 & 13.14 \\
   \hline
\end{tabular}
\caption{Clusters sizes and centers for $\Tau = 7.5$ minutes.}
\label{t:kmeans_centers_450}
\end{table*}

\begin{table*}[t]
\centering
\small
\begin{tabular}{c c c c c c c c c c}
  \hline
 Cluster &  Size & \thead{Total\\Trips} & \thead{Average\\Trips} & \thead{Average\\Length} & \thead{Average\\Distinct\\Origins} & \thead{Average\\Distinct\\Destinations} & \thead{Average\\First\\Hour} & \thead{Average\\Last\\Hour} & \thead{Average\\Rest\\Time} \\
  \hline
  1 & 371318 & 7.19 & 1.76 & 1.68 & 1.68 & 0.71 & 12.30 & 14.55 & 2.14 \\
    2 & 49852 & 44.85 & 3.73 & 1.90 & 3.16 & 1.73 & 8.98 & 17.02 & 7.76 \\
    3 & 189224 & 22.82 & 2.29 & 1.84 & 2.11 & 1.06 & 10.00 & 15.76 & 5.60 \\
    4 & 4593 & 114.34 & 9.12 & 2.30 & 5.92 & 3.98 & 6.93 & 18.21 & 10.43 \\
   \hline
\end{tabular}
\caption{Clusters sizes and  centers for $\Tau = 20$ minutes.}
\label{t:kmeans_centers_1200}
\end{table*}

### 2018-01-31 discussion.tex
Renamed section
Added conclusion and future work

% !TEX root = paper.tex
\section{Discussion}\label{s.conclusions}

>>>>

% !TEX root = paper.tex
\section{Conclusion and Future Work}\label{s.conclusions}

Most urban cities in the world possess a network of ANPR cameras that is used in law enforcement and for traffic monitoring and control. Number plate data collected in the Tyne and Wear area is stored and leveraged by the UTMC for computing average journey times across a selection of sensitive roads. However, number plate data could be used more extensively to identify and study individual and collective travel patterns. In this paper, we've presented a set of definitions and constraints that establish a conceptual foundation for identifying vehicle trips from number plate detections. We've identified two parameters, $\tau$ and $\Tau$ as critical in the discrimination of plausible and unplausible trips. Hence, future work should first and foremost focus on developing formal methods to estimate these parameters from observed distributions of travel times and by applying knowledge about the structure of the road network. Moreover, methods for addressing issues concerning camera performance, namely wrong and duplicate scans, should be further developed and researched.

Once trip data has been computed, a range of interesting applications is available. This work tries to identify groups of vehicles by clustering information about frequency and diversity of travel. By associating a vehicle with a cluster that represents taxis, or home-to-work commuters, one can begin estimating trip mode usage across the city. However, the results presented here could benefit from extra work and further validation. Finally, future work can focus on using trip data to solve interesting research problems such as: 
\begin{enumerate*}[label=(\roman*)]
  \item real-time route recommendation using probabilistic graphical models;
  \item detection of abnormal trip patterns for helping law enforcement in the identification of suspect vehicles or behavior;
  \item modelling how drivers make routing choices in the presence of anomalies in the road network.
\end{enumerate*}
